---
title: "Galaxy_Plots_on_AdaptreeEnvi_1R_CalcCorrs"
author: "Katie Lotterhos"
date: "1/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(!("ggplot2" %in% installed.packages())){install.packages("ggplot2")}
if(!("gstat" %in% installed.packages())){install.packages("gstat")}
if(!("sp" %in% installed.packages())){install.packages("sp")}   
if(!("maptools" %in% installed.packages())){install.packages("maptools")} 
if(!("ncf" %in% installed.packages())){install.packages("ncf")}   

source("https://bioconductor.org/biocLite.R")
biocLite("graph", suppressUpdates = TRUE)   
biocLite("Rgraphviz", suppressUpdates = TRUE)
biocLite("RBGL", suppressUpdates = TRUE)
install.packages("gRain", dependencies = TRUE)
install.packages("bnlearn")

library(ggplot2)
library(gstat)
library(sp)
library(maptools)
library(ncf)


library(graph)
library(Rgraphviz)
library(RBGL)
library(gRain)
library(bnlearn)
```

# example bayesian network
```{r}
a<- matrix(sample(c(1, 0), 15, TRUE), nrow=5, ncol=3)
bn_df <- data.frame(a)
head(bn_df)
res <- hc(bn_df)
res
plot(res)
```


setwd("/Users/katie/Desktop/CurrResearch/1-AdaptreeData/201509_PEGA/simulations/")

### Load data
```{r}
### envi
env <- read.table("data/results_AdaptreeEnviFor_R90.txt", header=TRUE)
names(env)
names(env)[13] <- "MAP"
names(env)[14] <- "MSP"

### Simulations
#snps <- read.table("data/LW2015data/1R_R90_1351142954_453_1_NumPops=90_NumInd=20.lfmm")
#env_sim <- read.table("data/LW2015data/1R_R90_1351142954_453_1_NumPops=90_NumInd=20.env")
#df <- read.table("data/LW2015data/1R_R90_1351142954_453_1_NumPops=90_NumInd=20Bayenv2LFMMpca.Cpval", header=TRUE)

snps1R <- read.table("data/LW2015data/1R_R90_1351142970_988_5_NumPops=90_NumInd=20.lfmm")
env_sim1R <- read.table("data/LW2015data/1R_R90_1351142970_988_5_NumPops=90_NumInd=20.env")
df1R <- read.table("data/LW2015data/1R_R90_1351142970_988_5_NumPops=90_NumInd=20Bayenv2LFMMpca.Cpval", header=TRUE)

snps2R <- read.table("data/LW2015data/2R_R90_1351142970_988_6_NumPops=90_NumInd=20.lfmm")
env_sim2R <- read.table("data/LW2015data/2R_R90_1351142970_988_6_NumPops=90_NumInd=20.env")
df2R <- read.table("data/LW2015data/2R_R90_1351142970_988_6_NumPops=90_NumInd=20Bayenv2LFMMpca.Cpval", header=TRUE)

identical(env_sim1R, env_sim2R) #good

env$SimEnv <- env_sim1R[seq(1,1800, by=20),1]

par(mfrow=c(1,1))
plot(env$Y_Pops, env$SimEnv)
  abline(lm(env$SimEnv~env$Y_Pops))
plot(env$Y_Pops, env$MAT)
  abline(lm(env$MAT~env$Y_Pops))
plot(env$X_Pops, env$MAT)
  abline(lm(env$MAT~env$X_Pops))

plot(env$X_Pops, env$SimEnv)
```


### Compare correlations with simulated environment
```{r}
head(env)
cor.with.sim <- cor(env$SimEnv, env[,c(2:3, 8:27)], method="spearman")
cor.with.sim
```

The correlations between the adaptree environment and the simulated environment are quite low.

### Calculate rho

Create a matrix to hold Spearman's rho and another matrix to hold $P$-values, where each row is a SNP and each column is an environmental variable

```{r}

### Create environment matrix for individual data
env2 <- matrix(NA, nrow(snps1R), 23) #hard coding
k=0
env_cols <- c(2:3, 8:28) #hard coding
for (i in env_cols){
  k=k+1
  env2[,k] <- rep(env[,i], each=20)
}
dim(env2)
colnames(env2) <- names(env)[env_cols]
head(env2)

### Create matrix for results
TrueEnv.rho1R <- TrueEnv.P1R <- matrix(NA, ncol(snps1R), ncol(env2))
TrueEnv.rho2R <- TrueEnv.P2R <- matrix(NA, ncol(snps2R), ncol(env2))
colnames(TrueEnv.rho1R) <- names(env)[env_cols]
colnames(TrueEnv.rho2R) <- names(env)[env_cols]
colnames(TrueEnv.P1R) <- names(env)[env_cols]
colnames(TrueEnv.P2R) <- names(env)[env_cols]

for (j in 1:ncol(env2)){
  print(colnames(env2)[j])
  for (i in 1:ncol(snps2R)){
    if(i<=ncol(snps1R)){
      test1R <- cor.test(snps1R[,i], env2[,j], method="spearman", exact=FALSE)
      TrueEnv.rho1R[i,j] <- test1R$estimate
      TrueEnv.P1R[i,j] <- test1R$p.value
    }
    
    test2R <- cor.test(snps2R[,i], env2[,j], method="spearman", exact=FALSE)
    TrueEnv.rho2R[i,j] <- test2R$estimate
    TrueEnv.P2R[i,j] <- test2R$p.value
  }
}
```

### Get indices of neutral and selected SNPs
(not shown)
```{r, echo=FALSE}
Neut1R = df1R$s_high==0 & df1R$UseSNP==TRUE
sum(Neut1R)
Neut2R = df2R$s_high==0 & df2R$UseSNP==TRUE
sum(Neut2R)

Sel1R = df1R$s_high>0 & df1R$UseSNP==TRUE
sum(Sel1R)
#which(Sel1R)
Sel2R = df2R$s_high>0 & df2R$UseSNP==TRUE
sum(Sel2R)
#which(Sel2R)

Neut_ind <- 1:9900
Sel_ind1R <- 9901:ncol(snps1R)
Sel_ind2R <- 9901:ncol(snps2R)
```


### How many are significant

I determined which loci were "outliers" by the criteria that they had to have p-values from Spearman's rho less than expected by Bonferroni correction.


#### False positives and true positives

```{r}
(Bonferroni_P_N <- -log10(0.05/(10000*22)))



### Number false positives in landscape sims
numFPPerEnv1R<-colSums(-log10(TrueEnv.P1R[Neut_ind,])>Bonferroni_P_N)
numFPPerEnv2R<-colSums(-log10(TrueEnv.P2R[Neut_ind,])>Bonferroni_P_N)


### Number true positives in landscape sims
numTPPerEnv1R <- colSums(-log10(TrueEnv.P1R[Sel_ind1R,])>Bonferroni_P_N)
  #(total selected loci in this dataset is 96)
numTPPerEnv2R <- colSums(-log10(TrueEnv.P2R[Sel_ind2R,])>Bonferroni_P_N, na.rm=TRUE)



cbind(numFPPerEnv1R, numFPPerEnv2R, numTPPerEnv1R, numTPPerEnv2R)
  # The false positive rate is really high in the simulated environment. Not sure why that is. It is true for the IBD sims, but not for the IM sims, which makes me think it has something to do with the scale of autocorrelation in the environment I simulated.

(order <- match(c("Y_Pops", "FFP", "MAP", "MWMT", "bFFP", "PAS", "X_Pops", "EMT", "MCMT", "TD", "DD0", "Elevation", "MAT", "eFFP", "NFFD", "DD5", "CMD", "Eref", "AHM", "MSP", "EXT", "SHM"), names(numFPPerEnv1R)))

```

### Clustering plots

```{r}
MakePlotDefault <- function(out,demog){
  sel_index <- length(which(out))-sum(which(out)>9900)+1 #index after which all under selection
  if(demog=="1R"){forh1R <- TrueEnv.rho1R[out, 1:22]}
  if(demog=="2R"){forh1R <- TrueEnv.rho2R[out, 1:22]}
  #tail(forh1R, 60)
  print(dim(forh1R))
  h1R <- hclust(dist(abs(forh1R)))
  #plot(h1R)
  
  # default clustering
  superheat(t(abs(forh1R)), 
          order.rows= rev(env_clust$rowInd),
          order.cols = h1R$order,
          heat.pal = (brewer.pal(5, "Blues")), heat.na.col = "white",
          col.dendrogram = T, row.dendrogram = TRUE,
          heat.lim = c(0,0.6),
          left.label.size = 0.45)
}

MakePlotOrdered <- function(out, demog){
  sel_index <- length(which(out))-sum(which(out)>9900)+1 #index after which all under selection
  if(demog=="1R"){forh1R <- TrueEnv.rho1R[out, 1:22]}
  if(demog=="2R"){forh1R <- TrueEnv.rho2R[out, 1:22]}
  #tail(forh1R, 60)
  print(dim(forh1R))
  h1R <- hclust(dist(abs(forh1R)))
  #plot(h1R)
  # superheat with columns ordered by selected or neutral
  superheat(t(abs(forh1R)), 
         order.rows= rev(env_clust$rowInd),
          order.cols = 1:nrow(forh1R),
          heat.pal = (brewer.pal(5, "Blues")), heat.na.col = "white",
          col.dendrogram = T, row.dendrogram = TRUE,
          heat.lim = c(0,0.6),
          left.label.size = 0.45)
}

### All 1R 
  out <- rowSums(-log10(TrueEnv.P1R[,1:22])>Bonferroni_P_N)>0
  MakePlotDefault(out, "1R")
  MakePlotOrdered(out, "1R")
### Neut 1R 
  out <- rowSums(-log10(TrueEnv.P1R[Neut_ind,1:22])>Bonferroni_P_N)>0
  MakePlotDefault(out, "1R")
### All 2R 
  out <- rowSums(-log10(TrueEnv.P2R[,1:22])>Bonferroni_P_N)>0
  MakePlotDefault(out, "2R")
  MakePlotOrdered(out, "2R")
### Neut 2R 
  out <- rowSums(-log10(TrueEnv.P2R[Neut_ind,1:22])>Bonferroni_P_N)>0
  MakePlotDefault(out, "2R")

```

```{r}
```

```{r}
### Barplot of FP vs. True Positives
  pdf(paste("results/1R_2R_FPvsTPBarplot_Bonf",round(Bonferroni_P_N ,2),".pdf", sep=""), width=7, height=8)
    par(mfrow=c(2,1), mar=c(6,4,0.5,0.5))
    barplot(rbind(c(numFPPerEnv1R[order]), numTPPerEnv1R[order]), las=2, ylab="Number of outlier loci", ylim=c(0,100))
    legend(17,90, c("# False positives", "# True positives"), fill=c("grey30", "grey70"), bty="n")
    text(0, 90, "A) One refuge model", adj=0)
    
    barplot(rbind(c(numFPPerEnv2R[order]), numTPPerEnv2R[order]), las=2, ylab="Number of outlier loci", ylim=c(0,100))
    text(0, 90, "B) Two refugia model", adj=0)
    dev.off()
```

```{r}
### Plot FP vs. TP as a function of correlation with the simulated environment
  pdf(paste("results/1R_2R_ScatterCorrWSimEnv_VsNumOut_Bonf",round(Bonferroni_P_N ,2),".pdf", sep=""), width=7, height=5)
  par(mfcol=c(2,2), mar=c(4,6,0.5,1), oma=c(0,0,3,0))
  plot(abs(cor.with.sim), numFPPerEnv1R[1:22], ylab="Number of false positives\nin focal environment", bty="l", xlab="", ylim=c(0,30))
  fplm <- lm(as.numeric(numFPPerEnv1R[1:22]) ~ as.numeric(abs(cor.with.sim)))
  summary(fplm)
  #abline(fplm)

  plot(abs(cor.with.sim), numTPPerEnv1R[1:22], ylab="Number of true positives\nin focal environment", bty="l", xlab="Correlation of focal environment\nwith selective environment", ylim=c(0,60))
    tplm <- lm(as.numeric(numTPPerEnv1R[1:22]) ~ as.numeric(abs(cor.with.sim)))
  summary(tplm)
  abline(tplm)
  text(0,)
  
   plot(abs(cor.with.sim), numFPPerEnv2R[1:22], ylab="Number of false positives\nin focal environment", bty="l", xlab="", ylim=c(0,30))
  fplm <- lm(as.numeric(numFPPerEnv2R[1:22]) ~ as.numeric(abs(cor.with.sim)))
  summary(fplm)
  #abline(fplm)

  plot(abs(cor.with.sim), numTPPerEnv2R[1:22], ylab="Number of true positives\nin focal environment", bty="l", xlab="Correlation of focal environment\nwith selective environment", ylim=c(0,60))
  tplm <- lm(as.numeric(numTPPerEnv2R[1:22]) ~ as.numeric(abs(cor.with.sim)))
  summary(tplm)
  abline(tplm)
  
  mtext("A) One refuge model", side=3, outer=TRUE, ad=0.25)
  mtext("B) Two refugia model", side=3, outer=TRUE, ad=0.87)
  dev.off()
```


The next step mimics what we did in the real data, which is to call a locus an outlier if it is an outlier in any one environmental variable.

```{r}
## 1R
  ### Get significance based on outlierliness in 22 measured variables (ignore unknown SimEnv)
  Is.Sig1R <- -log10(TrueEnv.P1R)>Bonferroni_P_N
  dim(Is.Sig1R)
  colnames(TrueEnv.P1R)
  colnames(Is.Sig1R)
  # As a first pass, we call a snp an outlier if it is an outlier in any one environmental variable (excluding the simulated environment)
  Is.Out1R <- (rowSums(Is.Sig1R[,1:22])>0)
  # The number of "outliers" that are neutral without knowing selective environment:
  sum(Is.Out1R[Neut_ind])
  # Number of "outliers" that are under selection without knowing selective environment:
  sum(Is.Out1R[Sel_ind1R])

## 2R
  Is.Sig2R <- -log10(TrueEnv.P2R)>Bonferroni_P_N
  Is.Out2R <- (rowSums(Is.Sig2R[,1:22])>0)
  sum(Is.Out2R[Neut_ind])
  sum(Is.Out2R[Sel_ind2R])
```


### Galaxy plots

```{r}
## 
MakePlot <- function(){
  Is.OutCol <- rep(adjustcolor("grey",0.2), length(Is.Out))
  Is.OutCol[which(Is.Out[Neut_ind] > 0)] <- adjustcolor("red", 0.5)
  Is.OutCol[Sel_ind] <- "grey70"
  Is.OutCol[Sel_ind[which(Is.Out[Sel_ind] > 0)]] <- "blue"
  Is.Outpch <- rep(19, length(Is.Out))
  Is.Outpch[Sel_ind] <- 25
  #Is.Outpch[Sel_ind[which(Is.Out[Sel_ind] > 0)]] <- 25
  #tail(cbind(Is.Outpch, Is.OutCol),100)
  
  TrueEnv.rho <- data.frame(TrueEnv.rho)
  TrueEnv.rho$Is.Out <- Is.Out
  TrueEnv.rho$Is.OutCol <- Is.OutCol
  TrueEnv.rho$s <- df$s_high[df$UseSNP]
  TrueEnv.rho$Is.Outpch <- Is.Outpch
  
  TrueEnv.rho2 <- TrueEnv.rho[order(TrueEnv.rho$s, TrueEnv.rho$Is.Out, decreasing=FALSE),]
  
  #par(mfrow=c(22,22), mar=c(0.5,0.5,0.5,0.5))
  
  names(TrueEnv.rho2)
  plotenv <- c(2,4, 8, 10:11)
  print(names(TrueEnv.rho2)[plotenv])
  ## Plot neutral only
#   png(paste("results/",nd,"_Paired_NeutOnly_Bonf",round(Bonferroni_P_N ,2),".png", sep=""), width=8, height=8, units = "in", res = 300)
#    pairs(TrueEnv.rho2[1:9900,c(plotenv,23)], pch=20, col=TrueEnv.rho2$Is.OutCol[1:9900])
#  dev.off()
    
  ## Plot neutral and selected loci with "signed" selected loci
  #png(paste("results/1R_Paired_Neut&SelSigned_Bonf",round(Bonferroni_P_N ,2),".png", sep=""), width=8, height=8, units = "in", res = 300)
   # pairs(TrueEnv.rho2[,c(plotenv,23)], pch=20, col=TrueEnv.rho2$Is.OutCol)
  #dev.off()
  
  # Randomly assign selected loci
  head(TrueEnv.rho2)
  TrueEnv.rho3 <- TrueEnv.rho2
  set.seed(1)
  flip <- sample(c(1,-1),size = length(Is.Out), replace=TRUE)
  for (i in 1:23){
    TrueEnv.rho3[,i] <- TrueEnv.rho3[,i] *flip
  }
  
  
  ## Plot neutral and selected loci with "random" selected loci
  png(paste("results/",nd,"_Paired_Neut&SelRandom_Bonf",round(Bonferroni_P_N ,2),".png", sep=""), width=8, height=8, units = "in", res = 300)
    pairs(TrueEnv.rho3[,c(plotenv,23)], pch=TrueEnv.rho3$Is.Outpch, col=TrueEnv.rho3$Is.OutCol)
    dev.off()
  
  ## Plot only selected loci
#  png(paste("results/1R_Paired_SelRandomOnly_Bonf",round(Bonferroni_P_N ,2),".png", sep=""), width=8, height=8, units = "in", res = 300)
#  cond <- TrueEnv.rho3$Is.Out==FALSE | TrueEnv.rho3$s>0
#    pairs(TrueEnv.rho3[cond,c(plotenv,23)], pch=20, col=TrueEnv.rho3$Is.OutCol[cond])
#    dev.off()
}# end plot function

Is.Out <- Is.Out1R
Sel_ind <- Sel_ind1R
TrueEnv.rho <- TrueEnv.rho1R
df <- df1R
nd <- "1R"
MakePlot()

Is.Out <- Is.Out2R
Sel_ind <- Sel_ind2R
TrueEnv.rho <- TrueEnv.rho2R
df <- df2R
nd <- "2R"
MakePlot()

```



```{r, eval=FALSE, echo=FALSE}
### Compare covariances and slopes
covdf <- matrix(NA, ncol=8, nrow=(sum(1:23)-24))
colnames(covdf) <- c("Env1", "Env2", "Genome_cov", "FP_cov", "TP_cov", "Genome_slope", "FP_slope", "TP_slope")
covdf <- data.frame(covdf)
FP_cond <- TrueEnv.rho3$s==0 & TrueEnv.rho3$Is.Out==TRUE
TP_cond <-  TrueEnv.rho3$s>0 & TrueEnv.rho3$Is.Out==TRUE

k=0
for (i in 1:21){
  for (j in (i+1):23){
    covdf[k,1] <- names(TrueEnv.rho3[i])
    covdf[k,2] <- names(TrueEnv.rho3[j])
    k=k+1
   covdf[k,3] <- as.numeric(cov(TrueEnv.rho3[,i], TrueEnv.rho3[,j]))
   
   covdf[k,4] <- as.numeric(cov(TrueEnv.rho3[FP_cond,i], TrueEnv.rho3[FP_cond,j]))
    covdf[k,5] <-  as.numeric(cov(TrueEnv.rho3[TP_cond,i], TrueEnv.rho3[TP_cond,j]))
    
    covdf[k,6] <- as.numeric(lm(TrueEnv.rho3[,i]~ TrueEnv.rho3[,j])$coeff[2])
    covdf[k,7] <- as.numeric(lm(TrueEnv.rho3[FP_cond,i]~ TrueEnv.rho3[FP_cond,j])$coeff[2])
    covdf[k,8] <- as.numeric(lm(TrueEnv.rho3[TP_cond,i]~ TrueEnv.rho3[TP_cond,j])$coeff[2])
    
  }
}


head(covdf)
tail(covdf, 50)
boxplot(log10(abs(covdf[,3:5])))
boxplot(abs(covdf[,6:8]))
```




```{r, eval=FALSE, echo=FALSE}
# trying some network models - doesn't work
Bonferroni_P_N <- 5
out1R <- -log10(TrueEnv.P1R[Sel_ind1R,1:23])>Bonferroni_P_N
colSums(out1R)
col <- colSums(out1R)>0
out1Rb <- apply(out1R,2, as.numeric)
out1Rc <- data.frame(out1Rb[,col])
tail(out1Rc, 50)
res <- hc(out1Rc)
res
quartz()
plot(res)
graphviz.plot(res)
```
